---
title: "Lab 4 - Potpourri"
categories: "Lab"
author: "Emily Bonanni"
date: "March 27, 2026"
format: "pdf"
---

```{r}
#| include: false
#| eval: true

library(datasauRus)
library(tidyverse)
library(palmerpenguins)
```

# Part 1 - All about Quarto

## Question 1

a. Add each of strings below to the code chunk provided in your document, render 
the document, and determine if the string is a proper code chunk label. If not, 
explain why and describe how you could fix it so the document renders.

- Chunk label 1 is not a proper chunk label because the line break does not 
allow it to render since the line break appear as a singular label. 

- Chunk label 2 is valid because it rendered and the label stays hidden in the 
output.

- Chunk label 3 did not work because labels cannot contain spaces. It did not 
render. 

- Chunk label 4 is valid since it was rendered using dashes.

b. Which of the chunk label options above is the best option? Explain your 
reasoning in 1-2 sentences.

Chunk label 4 is the best option because it is short and can allow us to use 
multiple words when using dashes. 

## Question 2

You have the following code chunk:

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```

Add the following code chunk options, one at a time, and set each to `false` and then to `true`. After each value, render your document and observe its effect. Ultimately, choose the values that are the most appropriate for this code chunk. Based on the behaviors you observe, describe what each code chunk option does.

-   `echo`
-   `warning`
-   `eval`

eval: If eval is true the code runs and output appears in document. If eval is
false then code is shown but never runs (no output).
echo: If echo is true then the code block is visible in the rendered document 
and the output below it is shown. If echo is false, the output is shown but the 
code is hidden
warning: If warning is true (if the code produces a warning), warning is shown 
in the rendered document. If warning is false, even if there is a warning it 
won't be shown in the rendered document.

## Question 3

a.  You have the following code chunk again.

```{r}
#| fig-width: 7
#| fig-asp: 0.7
ggplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point()
```
fig width of 7 and a fig aspect of 0.7 seems to look the best when rendering the pdf. fig width controls the width of the plot with higher values making it wider. fig aspect controls the height, where an aspect of 1 would make it a perfect square.

b.  You have the following code chunk, but look carefully, it's not exactly the same!

```{r}
#| error: true
gplot(penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
geom_point()
```
When typing error: true, the document will render and the code will show. 
When typing error: false, the document will not render at all.

# Part 2 - Misrepresentation

## Question 4

a.  What is misleading about this graph?

The visual is skewed as the number add to 100% but the bar length does not represent the percentages accurately. For example, the 87% bar should not be that much shorter than the 94% bar. It makes it appear that support for "private sector" and "don't know" have more support than they actually do. For example, for energy, the bars for private and don't know span more than half the chart but their combined percentages are only 13% which is very misleading.

b.  Suppose you wanted to recreate this plot, with improvements to avoid its misleading pitfalls from part (a). You would obviously need the data from the survey in order to be able to do that. How many observations would this data have? How many variables (at least) should it have, and what should those variables be?

The data has 1,858 observations total. There would be three variables if creating a new plot. The first variable would be the respondent (i.e. 1,2,3... 1,858), the second variable would be the service (rail, water, energy, and royal mail), and the last variable would be response (public sector, private sector, don't know). 

c.  Load the data for this survey from `data/survation.csv`. Confirm that the data match the percentages from the visualization. That is, calculate the percentages of public sector, private sector, don't know for each of the services and check that they match the percentages from the plot.

Data not provided. Unable to complete question. 

## Question 5

Data not provided. Unable to complete question.

# Part 3 - DatasauRus

The data frame you will be working with in this part is called `datasaurus_dozen` and it's in the **datasauRus** package. This single data frame contains 13 datasets, designed to show us why data visualization is important and how summary statistics alone can be misleading. The different datasets are marked by the `dataset` variable, as shown in @fig-datasaurus.


Here is a peek at the top 10 rows of the dataset:

```{r}
#| eval: true

datasaurus_dozen
```

## Question 6

In a single pipeline, calculate the mean of `x`, mean of `y`, standard deviation of `x`, standard deviation of `y`, and the correlation between `x` and `y` for each level of the `dataset` variable. Then, in 1-2 sentences, comment on how these summary statistics compare across groups (datasets).

```{r}
datasaurus_dozen %>%
  group_by(dataset) %>%
  summarize(
    mean_x=mean(x), 
    mean_y=mean(y), 
    s.d._x=sd(x), 
    s.d._y=sd(y), 
    correlation=cor(x,y)
  ) %>% print(n=13)
```

All of the datasets have the exact same means for x and y, and have the exact same standard deviations for x and y. The correlation values are near zero indicating no linear relationship between x and y, and are slightly different because of how x and y are paired together.

## Question 7

Create a scatterplot of `y` versus `x` and color and facet it by `dataset`. Then, in 1-2 sentences, how these plots compare across groups (datasets). How does your response in this question compare to your response to the previous question and what does this say about using visualizations and summary statistics when getting to know a dataset?

```{r}
ggplot(datasaurus_dozen, aes(x=x,y=y,colour=dataset)) +
  geom_point(show.legend = FALSE) + 
  facet_wrap(~dataset)
```
Although the summary statistics are nearly identical across all datasets, they look completely different when plotted. This shows that summary statistics can look the same but can be misleading because after plotting, the data can look quite different. 
